Heuristics for the Score-Constrained Bin-Packing Problem

### 2018-01-07 paper/paper.tex
changed break statements in algorithm ffdapprox
edited ffdapprox, started ffdexact

The one-dimensional bin packing problem (BPP) is a combinatorial optimisation problem that has been widely researched and discussed due to its ability to model a variety of real-life sitations.
The BPP can be described as follows: given a set of $n$ items of varying sizes $w$, and a finite number of bins with equal capacity $c$, find the minimum number of bins required to contain all of the items. As shown by \cite{garey1979}, this problem is NP-hard, and therefore (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution in polynomial time for every instance of the BPP. Instead, heursitics can be used to find a near-optimal solution in a shorter amount of time. One example is the greedy heuristic known as first-fit, an online algorithm that places each item (in some arbitrary order) into the lowest-indexed bin. If there are no bins that can feasibly accommodate the item, the algorithm places the item into a new bin. A modified version of first-fit yields the well-known first-fit decreasing (FFD) algorithm (\citealp{eilon1971}), which initially sorts the items into non-decreasing order of size before allocating them into bins. As proven by \cite{dosa2007}, the worst case for FFD is $\frac{11}{9}k + 4$ bins, where $k$ is the optimal number of bins given by $\ceil{\frac{\sum_{i=1}^{n} w_{i}}{c}}$ (..if the solution has this many bins then it is optimal, (\citealp{korf2002})).

@@

This leads us to our main problem: given a collection of boxes with varying widths and score widths on either side, and strips of a fixed length, find the minimum number of strips required to accommodate all of the boxes, such that the minimum total width constraint is met between all boxes on all strips.

@@

\begin{itemize}
	\item FF if online, FFD is offline
	\item FF and FFD can produce a feasible solution quickly, but the solution may not be optimal O(nln n), where n is the number of items that need to be packed
	\item Exact algorithms, Martello and Toth (MTP), Bin Completion algorithm Korf 2002, Schreiber and Korf (2013) (wiki)
	\item \textcolor{blue}{Although there are exact solutions available for the BPP} (martello knapsack MTP 1989 (1990b) ``is based on a first-fit decreasing'' branching strategy and is the best existing algorithm for optimal packing (KOrf2002)'', hung and brown 1978 ``a branch-and-bound algorithm'' paper title:``an algorithm for a class of loading problems'', eilon christofides 1971 ``a depth-first numerative algorithm'' paper title: ``the loadings problem'', schreiber and korf 2002 bin completion algorithm) \textcolor{blue}{they are only able to find solutions to problems that have a small input size.}
	\item ``Hence, exact solution techniques are bound to work well for small to medium sized problem instances only, and real world sized problems including up to thousands of items have to be solved heuristically.''
	\item ``There exists at least one ordering of items that allows FF to produce an optimal solution'' (wiki) Rhyd Lewis 2009
	\item Picture of box with score widths, score lines
	\item All boxes have same height, different widths, different score widths
	\item Recognition algorithm Becker
	\item Formal statement of BPP with score constraint called the Score Constrained Bin packing problem (SCBPP): given
	\item dicuss other paper/articles with this problem, i.e. rhyd 2017

\end{itemize}

\section{FFD Using Approximate Algorithm}
\label{sec:ffdapprox}

\begin{itemize}
	\item discuss heuristics
	\item ``Heuristics are approximate solution techniques that typically provide solutions in the least amount of time to the detriment of the solutions quality''
	\item ``Exact methods find a best possible packing, but are slow and may be unable to provide solutions to large or reastically sized problem instances within reasonable time''
	\item ``Consider a set of $n$ items. In the context of bin packing problems, heuristics consider these items one-by-one and attempt to pack them into a bin so as to achieve a near-optimal solution without the guarantee that a solution is optimal.''
	\item ``Exact packing methods are often called deterministic methods and are guarantees to find an optimal solution to a problem (given sufficient time and a feasible region) while heuristics attempt to find optimal solutions, but are not guaranteed to do so.''
	\item ``Exact methods may often be too slow to solve large problem instances, but may typically be used to solve smaller ones.''
	\item ``Packing heuristics follow a fixed set of rules to pack items in such a manner as to find good, feasible (but not necessarily optimal) solutions to the bin packing problem within as short a time span as possible.''
	\item offline algorithm, whole problem data is provided/required at the start.
	\item online, doesn't consider future events/items, data is processed in the order that it is input into the algorithm, entire input is not required to be available from the start
	\item ``first-fit decreasing (FFD) algorithm for one-dimensional bin packing by johnson et al (1974, worst-case performance bounds for simple one-dimensional packing algorithms)''
	\item FF is online, FFD is offline
	\item FF and FFD can produce a feasible solution quickly, but the solution may not be optimal, O(nln n), where n is the number of items that need to be packed
	\item who stated that ffd is O(nlog n)?
	\item Coffman et al 1984 heuristics survey
	\item \textcolor{blue}{Describe how the ffdapprox algorithm works}
	\item \textcolor{red}{figure of packed boxes on strips}
	\item previously placed boxes are fixed, cannot be rearranged
	\item complexity of modified FFD for SCBPP
	\item background information on FFD
	\item works in the same way as FFD, but with an additional step - each box must meet the minimum total score width constraint, so if a box fits onto a strip, we start by checking if the smallest of the two score widths on the box, and the free score width on the end of the current strip, meet the mtswc. If not, then we rotate the box $180^{\circ}$ and check again if the mtswc is met. If neither score width on the box meets the mtswc with the free score width, then we move onto the next strip and try again.
	\item start from bottom of strip/end of strip, pack from left to right
	\item Karp complexity
	\item \textcolor{red}{figure of one packed strip indicating free score end $e$}
	\item pros and cons: speed, however there may be a solution that exists that cannot be found due to boxes being fixed in place
	\item In this algorithm, the boxes are sorted by non-increasing widths prior to packing.
	\item If a box does not fit onto any existing strips, a new strip is initialised.
	\item Initially, pack the largest (widest/box with the largest width/first box in the sorted list of boxes) box into the first strip.

\end{itemize}

\begin{algorithm}[H]
	\caption{First-Fit Decreasing Algorithm for the Score-Constrained Bin-Packing Problem}
	\label{alg:ffdapprox}
	\begin{algorithmic}[1]
		\Require $n$ boxes with widths $w_i$, each box has two score widths $u_{i1} \text{ and } u_{i2}$, where $u_{i1}$ is the smaller of the two score widths, n strips of fixed lengths $l \text { and residual length } r_j \text { which is initially set to } l$. $j = 1,2, ..$ strips. $e_j$ is the score width at the end of strip $j$, i.e a score width of the most recent box packed into that strip, free end, etc. $t$ is the minimum total score width that must be met in order for the boxes to be packed. Assume no box has a width that is larger than the strip length, i.e. $w_i \leq l \forall i $. Could do strips set $S$ with $\{u_{i1}, u_{i2}\}$ representing box on strip $S_j$ with smallest score width first, and $\{u_{i2}, u_{i1}\}$ representing box with larger score width first. need to put $i = 1, 2, ..., n$ and $j = 1, 2, ...$.
		\State Sort boxes in non-increasing order of widths, breaking ties arbitrarily
		\ForAll{boxes $i = 1, 2, ..., n$}
			\ForAll{strips $j = 1,2,...$}
				\If{$r_j < l \land w_i \leq r_j $}
					\If{$e_j + s_{i1} \geq t$}
						\IndIndIndStatex $S_j \gets S_j + \{\{u_{i1}, u_{i2}\}\}$
						\IndIndIndStatex $r_j \gets r_j - w_i$
						\IndIndIndStatex break
					\ElsIf{$e_j + s_{i2} \geq t$}
						\IndIndIndStatex $S_j \gets S_j + \{\{u_{i2}, u_{i1}\}\}$
						\IndIndIndStatex $r_j \gets r_j - w_i$
						\IndIndIndStatex break
					\EndIf
				\ElsIf{$r_j = l$}
					\IndIndStatex $S_j \gets S_j + \{\{u_{i1}, u_{i2}\}\}$
					\IndIndStatex $r_j \gets r_j - w_i$
					\IndIndStatex break
				\EndIf
			\EndFor
		\EndFor
		\Ensure all boxes packed onto strips while meeting minimum total width constraint. Set $S$ is all strips with boxes in correct rotation to meet minimum total score width constraint.

	\end{algorithmic}	
\end{algorithm}

\section{FFD Using Exact Algorithm}
\label{sec:ffdexact}

@@

>>>>

The one-dimensional bin packing problem (BPP) is a combinatorial optimisation problem that has been widely researched and discussed due to its ability to model a variety of real-life sitations.
The BPP can be described as follows: given a set of $n$ items of varying sizes $w$, and a finite number of bins with equal capacity $c$, find the minimum number of bins required to contain all of the items. As shown by \cite{garey1979}, this problem is NP-hard, and therefore (under the assumption that $P \neq NP$) there is no known algorithm that is able to find an optimal solution in polynomial time for every instance of the BPP. Instead, heursitics can be used to find a near-optimal solution in a shorter amount of time. One example is the greedy heuristic known as first-fit, an online algorithm that places each item (in some arbitrary order) into the lowest-indexed bin. If there are no bins that can feasibly accommodate the item, the algorithm places the item into a new bin. A modified version of first-fit yields the well-known first-fit decreasing (FFD) algorithm (\citealp{eilon1971}), which initially sorts the items into non-decreasing order of size before allocating them into bins. As proven by \cite{dosa2007} (tight bound), the worst case for FFD is $\frac{11}{9}k + \frac{6}{9}$ bins, where $k$ is the optimal number of bins given by $\ceil{\frac{\sum_{i=1}^{n} w_{i}}{c}}$ (..if the solution has this many bins then it is optimal, (\citealp{korf2002})).

@@

This leads us to our main problem, the score-constrained bin-packing problem: given a collection of boxes with varying widths and score widths on either side, and strips of a fixed length, find the minimum number of strips required to accommodate all of the boxes, such that the minimum total width constraint is met between all boxes on all strips.

@@

\begin{itemize}
	\item FF if online, FFD is offline
	\item FF and FFD can produce a feasible solution quickly, but the solution may not be optimal O(nln n), where n is the number of items that need to be packed
	\item Exact algorithms, Martello and Toth (MTP), Bin Completion algorithm Korf 2002, Schreiber and Korf (2013) (wiki)
	\item \textcolor{blue}{Although there are exact solutions available for the BPP} (martello knapsack MTP 1989 (1990b) ``is based on a first-fit decreasing'' branching strategy and is the best existing algorithm for optimal packing (KOrf2002)'', hung and brown 1978 ``a branch-and-bound algorithm'' paper title:``an algorithm for a class of loading problems'', eilon christofides 1971 ``a depth-first numerative algorithm'' paper title: ``the loadings problem'', schreiber and korf 2002 bin completion algorithm) \textcolor{blue}{they are only able to find solutions to problems that have a small input size.}
	\item ``Hence, exact solution techniques are bound to work well for small to medium sized problem instances only, and real world sized problems including up to thousands of items have to be solved heuristically.''
	\item ``There exists at least one ordering of items that allows FF to produce an optimal solution'' (wiki) Rhyd Lewis 2009
	\item Picture of box with score widths, score lines
	\item All boxes have same height, different widths, different score widths
	\item Recognition algorithm Becker
	\item Formal statement of BPP with score constraint called the Score Constrained Bin packing problem (SCBPP): given
	\item dicuss other paper/articles with this problem, i.e. rhyd 2017	
\end{itemize}

\section{Formulating the problem}
We are given $n$ boxes that need to be packed onto strips. Each of these boxes have a positive width, and two score widths, one on either side of the box. Since these score widths are not necessarily equal, the boxes can be aligned in two ways: either in a ``regular'' alignment, where the smaller of the two score widths is on the left hand side of the box, or in a ``rotated'' alignment, where the box has been rotated by $180^{\circ}$ so that the larger score width is on the left-hand side of the box. Strips are packed left to right. We also have a finite number of strips of equal fixed lengths. Once a box has been placed onto a strip, the rightmost score on the strip is referred to as the ``free'' score. A box can only be placed onto a strip if it can fit onto the strip, i.e. the addition of the box to the strip doesn't exceed the length of the strip, and if the free score and one of the scores on the box meet the minimum total score width constraint, i.e. the sum of the free score width and one of the scores on the box is greater than or equal to the minimum total score width. Assume that no box has a width that is greater than the fixed length of the strips.

Definition: Score-Constrained Bin-Packing Problem. Find the minimum number of strips required to feasibly pack all the boxes such that the minimum total score width constraint is met between every pair of adjacent score widths from different boxes on every strip. Then the SCBPP consists of finding a feasible alignment of all of the boxes onto the strips such that every pair of adjacent score widths from different boxes meets the minimum total score width constraint, and that the least number of strips is used.

\section{FFD Using Approximate Algorithm}
\label{sec:ffdapprox}
\begin{itemize}
	\item discuss heuristics
	\item ``Heuristics are approximate solution techniques that typically provide solutions in the least amount of time to the detriment of the solutions quality''
	\item ``Exact methods find a best possible packing, but are slow and may be unable to provide solutions to large or reastically sized problem instances within reasonable time''
	\item ``Consider a set of $n$ items. In the context of bin packing problems, heuristics consider these items one-by-one and attempt to pack them into a bin so as to achieve a near-optimal solution without the guarantee that a solution is optimal.''
	\item ``Exact packing methods are often called deterministic methods and are guarantees to find an optimal solution to a problem (given sufficient time and a feasible region) while heuristics attempt to find optimal solutions, but are not guaranteed to do so.''
	\item ``Exact methods may often be too slow to solve large problem instances, but may typically be used to solve smaller ones.''
	\item ``Packing heuristics follow a fixed set of rules to pack items in such a manner as to find good, feasible (but not necessarily optimal) solutions to the bin packing problem within as short a time span as possible.''
	\item offline algorithm, whole problem data is provided/required at the start.
	\item online, doesn't consider future events/items, data is processed in the order that it is input into the algorithm, entire input is not required to be available from the start
	\item ``first-fit decreasing (FFD) algorithm for one-dimensional bin packing by johnson et al (1974, worst-case performance bounds for simple one-dimensional packing algorithms)''
	\item FF is online, FFD is offline
	\item FF and FFD can produce a feasible solution quickly, but the solution may not be optimal, O(nln n), where n is the number of items that need to be packed
	\item who stated that ffd is O(nlog n)?
	\item Coffman et al 1984 heuristics survey
	\item \textcolor{blue}{Describe how the ffdapprox algorithm works}
	\item \textcolor{red}{figure of packed boxes on strips}
	\item previously placed boxes are fixed, cannot be rearranged
	\item complexity of modified FFD for SCBPP
	\item background information on FFD
	\item works in the same way as FFD, but with an additional step - each box must meet the minimum total score width constraint, so if a box fits onto a strip, we start by checking if the smallest of the two score widths on the box, and the free score width on the end of the current strip, meet the mtswc. If not, then we rotate the box $180^{\circ}$ and check again if the mtswc is met. If neither score width on the box meets the mtswc with the free score width, then we move onto the next strip and try again.
	\item pack from left to right
	\item boxes can be aligned in two ways, with smaller score on LHS, or rotated $180^{\circ}$ so the smaller score is on the LHS
	\item Karp complexity
	\item \textcolor{red}{figure of one packed strip indicating free score end $e$}
	\item pros and cons: speed, however there may be a solution that exists that cannot be found due to boxes being fixed in place
	\item In this algorithm, the boxes are sorted by non-increasing widths prior to packing.
	\item If a box does not fit onto any existing strips, a new strip is initialised.
	\item Initially, pack the largest (widest/box with the largest width/first box in the sorted list of boxes) box into the first strip.
	\item ``Simchi-Levi (1994, ``New worst-case results for the bpp'') has shown that the FFD and the BFD heuristics have an absolute performance ratio of 1.5, and that this value is the best possible for the BPP, unles $P = NP$.''
\end{itemize}

\begin{algorithm}[H]
\caption{First-Fit Decreasing Algorithm for the Score-Constrained Bin-Packing Problem}
\label{alg:ffdapprox}
	\begin{algorithmic}[1]
	\Require $n$ boxes with widths $w(i)$ for $i = 1,2,...,n$, each box $i$ has two scores $i_1$ (smaller score width) and $i_2$ (larger score width), where $w(i_1) \leq w(i_2)$. Let $\{i_1, i_2\}$ represent box $i$ aligned such that the smaller score is on the LHS, and let $\{i_2, i_2\}$ represent box $i$ rotated by $180^{\circ}$ such that the larger score is on the LHS. Strips are packed from left to right. Finite number of strips $S_j, j = 1, 2, ...$ that all have equal fixed length $l > 0$, and a residual length $r(j)$, where initially $r(j) = l \forall j = 1, 2, ...$. Initially $S_j = \emptyset \forall j = 1, 2,...$. Let $w(f_j)$ denote the width of the free score for strip $j$.   
	\State Sort boxes in non-increasing order of widths, breaking ties arbitrarily
	\State $i \gets 1, j \gets 1$
	\State $S_j \gets S_j + \{\{i_1, i_2\}\}$
	\State $r(j) \gets r(j) - w(i)$
	\For{$i \gets 2$ \textbf{to} $n$}
		\For{$j \gets 1$ \textbf{to} $n$}
			\If{$S_j \neq \emptyset \land w(i) \leq r(j)$}
				\If{$w(f_j) + w(i_1) \geq t$}
					\IndIndIndStatex $S_j \gets S_j \cup \{i_1, i_2\}$
					\IndIndIndStatex $r(j) \gets r(j) - w(i)$
					\IndIndIndStatex \Break
				\ElsIf{$w(f_j) + w(i_2) \geq t$}
					\IndIndIndStatex $S_j \gets S_j \cup \{i_2, i_1\}$
					\IndIndIndStatex $r(j) \gets r(j) - w(i)$
					\IndIndIndStatex \Break					
				\EndIf
			\ElsIf{$S_j = \emptyset$}
				\IndIndStatex $S_j \gets S_j \cup \{i_1, i_2\}$
				\IndIndStatex $r(j) \gets r(j) - w(i)$
				\IndIndStatex \Break
			\EndIf
		\EndFor
	\EndFor		
	\Ensure all boxes packed onto strips while meeting minimum total width constraint. Set $S$ is all strips with boxes in correct rotation to meet minimum total score width constraint.	
	\end{algorithmic}	
\end{algorithm}

DESCRIPTION OF FFDAPPROX 
\begin{itemize}
	\item start by sorting all boxes into non-increasing order of widths, breaking ties arbitrarily
	\item starting with the first box in the list (i.e with the box that has the largest width), put it onto the first strip, with the box rotated such that the side of the box with the smaller score width is placed at the beginning of the strip
	\item for each box, find the lowest indexed strip that the box can fit into
	\item if the box can fit onto a strip, check if the free score width on that strip and the smaller of the two score widths on the box meet the minimum total score width requirement. If so, place the box onto that strip such that the smaller of the two score widths is placed next to the free score width. Move onto the next box in the list
	\item else, check if the larger of the two score widths on the box and the free score width meet the minimum total score width constraint. If so, place the box onto that strip such that the larger of the two score widths is placed next to the free score width. Move onto the next box.
	\item else, if neither of the two score widths and the free score meet the minimum total score width constraint, move onto the next used strip and repeat
	\item if there are no used strips that can accommodate the current box, place the box into a new unused strip, with the smaller score at the beginning of the strip (on the LHS). move onto the next box.
	\item we then obtain ...
\end{itemize}

\section{FFD Using Exact Algorithm}
\label{sec:ffdexact}
Introduction
\begin{itemize}
	\item This alg will be combining a well known heuristic with an exact algorithm to find a feasible solution to the SCBPP
	\item Becker 2010 recognicity algorithm in polynomial time
	\item doesn't take into account individual widths of boxes
	\item only one strip/align all boxes into one row
	\item in industry, strips have a finite length
	\item algorithm also doesn't output a full solution, only whether there exists a feasible solution or not
	\item ffd but including an algorithm that can find a feasible aligment of boxes with score constraints if one exists
	\item unlike ffd, where boxes are placed onto the end of each strip, the boxes are not fixed in place, all boxes can be reordered.
	\item try to find a hamiltonian path 
\end{itemize}

Explanation: Initialise
\begin{itemize}
	\item initialise
\end{itemize}

Let each box be represented by two vertices on a graph $G$, one vertex for each side of the box. Each vertex has a value corresponding to the score width it represent. We will refer to two vertices that represent either side of the same box as ``mates''. The boxes cannot be altered in any way, and so the pairs of vertices must be adjacent to one another on the graph.

------------------

MTGMA/MIS
\begin{itemize}
	\item mtgma/mis
\end{itemize}

FIS/Patch
\begin{itemize}
	\item fis/patch
\end{itemize}

@@

%%Additional Comments
\begin{comment}
- all citations have been changed to \cite{}, so that the citation shows up in green in latex. some of these citations need to be changed to (\citealp{}) (or \citep{}, same outcome) as the citations need to be in brackets but without the year in brackets.
- 1. "As proven by \cite{dosa2007}, ..." -> "As proven by Dosa (2007), ..."
- 2. "... FFD algorithm (\cite{eilon1971}), which ..." -> "... FFD algorithm (Eilon and Christofides (1971)), which .." NOTE THE YEAR IS IN BRACKETS INSIDE THE CITATION BRACKETS, THIS IS ONLY WHILE WRITING, THIS NEEDS TO BE CHANGED BEFORE SUBMISSION
- 3. "... FFD algorithm (\citealp{eilon1971}), which ..." -> "... FFD algorithm (Eilon and Christofides, 1971), which .." THIS IS HOW IT NEEDS TO BE WHEN SUBMITTING
\end{comment}
